Writing this dissertation has been a long, difficult, amazing, exhausting, and ultimately rewarding task.  I am grateful to have so many people to thank for their support along the way.

To my advisors: Michael Friedlander, thank for inspiring me with your engaging courses and supporting my research in my early graduate school years.  My passion for optimization and numerical methods was shaped in no small part by your advising.  
And to Zhaojun Bai, special thanks for your support, commitment, and time spent over the past three years.  When I was a fourth-year graduate student, you offered to advise me throughout the entire dissertation research and writing process. 
Your guidance helped me see the forest for the trees, keeping my eyes on the overall goals of this dissertation.  Along the way, you taught me how to identify important research topics, organize research papers, and produce quality technical writing.  I truly could not have done this without you.  Thank you.

To my family: Thank you all for believing in me and supporting me throughout my graduate work.  To my mother Sharon, my daughter Julia, and her mother Pattie, thank you all for listening to my challenges and offering encouraging words.  I could not count the number of phone calls I had with you, discussing the latest challenge or worrying over the direction of my research.  Your support meant everything to me.

And to my friends: Thank you all for your support and understanding.  To my Davis friends from the \textit{Anderson House}: our evenings spent discussing the day, watching movies, and playing games like \textit{Pandemic Legacy: Season 2} and \textit{Overcooked!} were exactly what I needed to maintain my sanity.
To my West Coast friends: thank you for being there whenever I visited you in Paradise, Portland, San Diego, or San Francisco.
Leaving the academic bubble of Davis to visit you was always a refreshing and rewarding change of pace.
And, for anyone who wants to ask for the $1,000,000\textsuperscript{th}$ time if my dissertation is done, the answer is ``yes''.


